---
title: "datawrangle_dietguthealth"
author: "Yasmine"
date: "2024-01-09"
output: html_document
---
**Purpose**: Wrangle data for analysis for USDA Nutritional Phenotyping Study associations between diet and gastrointestinal inflammation project. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries 
library(tidyverse)
```

Variables of interest (ASA24 and FFQ): 
- Energy
- Fiber
- Vegetables (excluding legumes)
- Legumes
- Total vegetables 
- Saturated fat
- HEI scores
- DII scores

Separate data frame for each GIM:
- Calprotectin (df_c)
- Myeloperoxidase (df_m)
- Neopterin (df_n) 
- LPS-binding protein (df_l)


```{r data-import}

# Import data frames with GI inflammation data 
# For stool markers (calprotectin, MPO, neopterin), stool collected post-V2 and stool processed >24h have been excluded

# For all markers, transformation to normal has been conducted for statistical tests: clpn_ln is (ln(x + 3)), mpo_ln is (ln(x + 1)), nptn_boxcox is Box Cox, lbp_ln is (ln(x + 1))

df_c <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/calprotectin_data.csv", header = T)

df_m <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/mpo_data.csv", header = T)

df_n <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/neopterin_data.csv", header = T)

df_l <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/lbp_data.csv", header = T)

```

```{r append}

# Append dietary data from ASA24

asa24 <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/forR_recall_data.csv", header = T)

df_c <- merge(df_c, asa24, by = 'subject_id')
df_m <- merge(df_m, asa24, by = 'subject_id')
df_n <- merge(df_n, asa24, by = 'subject_id')
df_l <- merge(df_l, asa24, by = 'subject_id')

# Append dietary data from FFQ

ffq <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/forR_ffq_data.csv", header = T)

df_c <- merge(df_c, ffq, by = 'subject_id')
df_m <- merge(df_m, ffq, by = 'subject_id')
df_n <- merge(df_n, ffq, by = 'subject_id')
df_l <- merge(df_l, ffq, by = 'subject_id')

# Append HEI scores

hei <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/forR_HEIscores.csv", header = T)

df_c <- merge(df_c, hei, by = 'subject_id')
df_m <- merge(df_m, hei, by = 'subject_id')
df_n <- merge(df_n, hei, by = 'subject_id')
df_l <- merge(df_l, hei, by = 'subject_id')

# Append DII scores 

dii <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/forR_DIIscores.csv", header = T)

df_c <- merge(df_c, dii, by = 'subject_id')
df_m <- merge(df_m, dii, by = 'subject_id')
df_n <- merge(df_n, dii, by = 'subject_id')
df_l <- merge(df_l, dii, by = 'subject_id')

```

```{r correct-variable-data-types}

df_c$subject_id <- as.factor(as.character(df_c$subject_id))
df_c$sex <- df_c$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_c$age_cat <- df_c$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_c$bmi_cat <- df_c$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_m$subject_id <- as.factor(as.character(df_m$subject_id))
df_m$sex <- df_m$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_m$age_cat <- df_m$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_m$bmi_cat <- df_m$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_n$subject_id <- as.factor(as.character(df_n$subject_id))
df_n$sex <- df_n$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_n$age_cat <- df_n$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_n$bmi_cat <- df_n$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_l$subject_id <- as.factor(as.character(df_l$subject_id))
df_l$sex <- df_l$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_l$age_cat <- df_l$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_l$bmi_cat <- df_l$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

rm(asa24, ffq, dii, hei)

# READY FOR LIFT OFF

```

For reproducibility, export each of these data frames into the 'data' folder
```{r export}

write.csv(df_c, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_c.csv", row.names = F)

write.csv(df_m, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_m.csv", row.names = F)

write.csv(df_n, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_n.csv", row.names = F)

write.csv(df_l, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_l.csv", row.names = F)

```

They will be loaded at the start of each Rmd file. 

