---
title: "table1_dietguthealth"
author: "Yasmine"
date: "2024-01-12"
output: html_document
---

Running statistics to see associations between fiber and gut health (Table 1)
Represent each variable as units per 1000 kcal for standardization 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

```{r import}

df_c <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_c.csv", header = T)

df_m <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_m.csv", header = T)

df_n <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_n.csv", header = T)

df_l <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_l.csv", header = T)

# unfortunately, need to correct structure of some variables each time -->

df_c$subject_id <- as.factor(as.character(df_c$subject_id))
df_c$sex <- df_c$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_c$age_cat <- df_c$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_c$bmi_cat <- df_c$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_m$subject_id <- as.factor(as.character(df_m$subject_id))
df_m$sex <- df_m$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_m$age_cat <- df_m$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_m$bmi_cat <- df_m$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_n$subject_id <- as.factor(as.character(df_n$subject_id))
df_n$sex <- df_n$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_n$age_cat <- df_n$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_n$bmi_cat <- df_n$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_l$subject_id <- as.factor(as.character(df_l$subject_id))
df_l$sex <- df_l$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_l$age_cat <- df_l$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_l$bmi_cat <- df_l$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

```

Models used for analysis: regression of transformed GI inflammation marker ~ dietary component
Covariates: age, sex, BMI

```{r asa24-total-fiber}

# Create fibe_pk (pk = per 1000 kcal)

df_c$fibe_pk <- df_c$fibe/(df_c$kcal/1000)
df_m$fibe_pk <- df_m$fibe/(df_m$kcal/1000)
df_n$fibe_pk <- df_n$fibe/(df_n$kcal/1000)
df_l$fibe_pk <- df_l$fibe/(df_l$kcal/1000)

# Calprotectin

mod_fibe_pk_c <- lm(clpn_ln ~ fibe_pk + age + sex + bmi_final, data = df_c)
summary(mod_fibe_pk_c) # 

# Myeloperoxidase

mod_fibe_pk_m <- lm(mpo_ln ~ fibe_pk + age + sex + bmi_final, data = df_m)
summary(mod_fibe_pk_m) # NS

# Neopterin

mod_fibe_pk_n <- lm(nptn_boxcox ~ fibe_pk + age + sex + bmi_final, data = df_n)
summary(mod_fibe_pk_n) # p = 0.00252 **

# LBP

mod_fibe_pk_l <- lm(lbp_ln ~ fibe_pk + age + sex + bmi_final, data = df_l)
summary(mod_fibe_pk_l) # NS

# Extract p-values 
p_c_fib <- summary(mod_fibe_pk_c)$coefficients[2,4] # 
p_m_fib <- summary(mod_fibe_pk_m)$coefficients[2,4]
p_n_fib <- summary(mod_fibe_pk_n)$coefficients[2,4]
p_l_fib <- summary(mod_fibe_pk_l)$coefficients[2,4]


```
Next, we'll look at total fiber from FFQ (dt_fibe)

```{r ffq-total-fiber}

# Create dt_fibe_pk (pk = per 1000 kcal)

df_c$dt_fibe_pk <- df_c$dt_fibe/(df_c$kcal/1000)
df_m$dt_fibe_pk <- df_m$dt_fibe/(df_m$kcal/1000)
df_n$dt_fibe_pk <- df_n$dt_fibe/(df_n$kcal/1000)
df_l$dt_fibe_pk <- df_l$dt_fibe/(df_l$kcal/1000)

# Calprotectin

mod_dt_fibe_pk_c <- lm(clpn_ln ~ dt_fibe_pk + age + sex + bmi_final, data = df_c)
summary(mod_dt_fibe_pk_c) # NS

# Myeloperoxidase

mod_dt_fibe_pk_m <- lm(mpo_ln ~ dt_fibe_pk + age + sex + bmi_final, data = df_m)
summary(mod_dt_fibe_pk_m) # NS

# Neopterin

mod_dt_fibe_pk_n <- lm(nptn_boxcox ~ dt_fibe_pk + age + sex + bmi_final, data = df_n)
summary(mod_dt_fibe_pk_n) # NS

# LBP

mod_dt_fibe_pk_l <- lm(lbp_ln ~ dt_fibe_pk + age + sex + bmi_final, data = df_l)
summary(mod_dt_fibe_pk_l) # NS

# Extract p-values (_f for ffq)
p_c_fib_f <- summary(mod_dt_fibe_pk_c)$coefficients[2,4] 
p_m_fib_f <- summary(mod_dt_fibe_pk_m)$coefficients[2,4]
p_n_fib_f <- summary(mod_dt_fibe_pk_n)$coefficients[2,4]
p_l_fib_f <- summary(mod_dt_fibe_pk_l)$coefficients[2,4]


```

Last, we'll look at soluble fiber from the FFQ (dt_fiber_sol)

```{r ffq-sol-fiber}


# Create dt_fiber_sol_pk (pk = per 1000 kcal)

df_c$dt_fiber_sol_pk <- df_c$dt_fiber_sol/(df_c$kcal/1000)
df_m$dt_fiber_sol_pk <- df_m$dt_fiber_sol/(df_m$kcal/1000)
df_n$dt_fiber_sol_pk <- df_n$dt_fiber_sol/(df_n$kcal/1000)
df_l$dt_fiber_sol_pk <- df_l$dt_fiber_sol/(df_l$kcal/1000)

# Calprotectin

mod_dt_fiber_sol_pk_c <- lm(clpn_ln ~ dt_fiber_sol_pk + age + sex + bmi_final, data = df_c)
summary(mod_dt_fiber_sol_pk_c) # NS

# Myeloperoxidase

mod_dt_fiber_sol_pk_m <- lm(mpo_ln ~ dt_fiber_sol_pk + age + sex + bmi_final, data = df_m)
summary(mod_dt_fiber_sol_pk_m) # NS

# Neopterin

mod_dt_fiber_sol_pk_n <- lm(nptn_boxcox ~ dt_fiber_sol_pk + age + sex + bmi_final, data = df_n)
summary(mod_dt_fiber_sol_pk_n) # NS

# LBP

mod_dt_fiber_sol_pk_l <- lm(lbp_ln ~ dt_fiber_sol_pk + age + sex + bmi_final, data = df_l)
summary(mod_dt_fiber_sol_pk_l) # NS

# Extract p-values (_f for ffq)
p_c_sfib_f <- summary(mod_dt_fiber_sol_pk_c)$coefficients[2,4] 
p_m_sfib_f <- summary(mod_dt_fiber_sol_pk_m)$coefficients[2,4]
p_n_sfib_f <- summary(mod_dt_fiber_sol_pk_n)$coefficients[2,4]
p_l_sfib_f <- summary(mod_dt_fiber_sol_pk_l)$coefficients[2,4]


```

```{r p-test-df}

# Collating the p-vals into a df for future multiple hypothesis test correction 

c_asa_pk_pval_df <- data.frame(p_c_fib)
c_ffq_pk_pval_df <- data.frame(p_c_fib_f, p_c_sfib_f)

m_asa_pk_pval_df <- data.frame(p_m_fib)
m_ffq_pk_pval_df <- data.frame(p_m_fib_f, p_m_sfib_f)

n_asa_pk_pval_df <- data.frame(p_n_fib)
n_ffq_pk_pval_df <- data.frame(p_n_fib_f, p_n_sfib_f)

l_asa_pk_pval_df <- data.frame(p_l_fib)
l_ffq_pk_pval_df <- data.frame(p_l_fib_f, p_l_sfib_f)

# Export

write.csv(c_asa_pk_pval_df, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/c_asa_pk_pval_df.csv", row.names = T)
write.csv(c_ffq_pk_pval_df, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/c_ffq_pk_pval_df.csv", row.names = T)

write.csv(m_asa_pk_pval_df, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/m_asa_pk_pval_df.csv", row.names = T)
write.csv(m_ffq_pk_pval_df, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/m_ffq_pk_pval_df.csv", row.names = T)

write.csv(n_asa_pk_pval_df, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/n_asa_pk_pval_df.csv", row.names = T)
write.csv(n_ffq_pk_pval_df, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/n_ffq_pk_pval_df.csv", row.names = T)

write.csv(l_asa_pk_pval_df, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/l_asa_pk_pval_df.csv", row.names = T)
write.csv(l_ffq_pk_pval_df, "/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/l_ffq_pk_pval_df.csv", row.names = T)

```

**Summary of Findings**: 

- 