---
title: "pa_corr_dietguthealth"
author: "Yasmine"
date: "2024-01-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggpubr)
library(rcompanion)
```

```{r data}

df_c <- read.csv("/Users/yasmine.bouzid/Desktop/GraduateWork/DataScience/Dairy_GutHealth/data/df_c_sc.csv", header = T) # n = 295, check. 

df_m <- read.csv("/Users/yasmine.bouzid/Desktop/GraduateWork/DataScience/Dairy_GutHealth/data/df_m_sc.csv", header = T) # n = 295, check. 

df_n <- read.csv("/Users/yasmine.bouzid/Desktop/GraduateWork/DataScience/Dairy_GutHealth/data/df_n_sc.csv", header = T) # 289, check. 

# and normalize:

df_n$nptn_boxcox <- (bestNormalize::boxcox(df_n$fecal_neopterin))$x.t

df_l <- read.csv(file = "/Users/yasmine.bouzid/Desktop/GraduateWork/DataScience/Dairy_GutHealth/data/df_gh_l_rd.csv", header = TRUE) #348, check. 

# And fix the structure of things 

df_c$subject_id <- as.factor(as.character(df_c$subject_id))
df_c$sex <- df_c$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_c$age_cat <- df_c$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_c$bmi_cat <- df_c$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")
df_c$AfterV2 <- df_c$AfterV2 %>% recode_factor('0' = "No", '1' = "Yes")

df_m$subject_id <- as.factor(as.character(df_m$subject_id))
df_m$sex <- df_m$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_m$age_cat <- df_m$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_m$bmi_cat <- df_m$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")
df_m$AfterV2 <- df_m$AfterV2 %>% recode_factor('0' = "No", '1' = "Yes")

df_n$subject_id <- as.factor(as.character(df_n$subject_id))
df_n$sex <- df_n$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_n$age_cat <- df_n$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_n$bmi_cat <- df_n$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")
df_n$AfterV2 <- df_n$AfterV2 %>% recode_factor('0' = "No", '1' = "Yes")

df_l$subject_id <- as.factor(as.character(df_l$subject_id))
df_l$sex <- df_l$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_l$age_cat <- df_l$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_l$bmi_cat <- df_l$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

# READY FOR LIFT OFF

```

Import the physical activity (PA) data and append 

```{r pa-data}

# Import

pa <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/forR_pa_data.csv", header = T)

# Append to the GIM dataframes 

df_c <- merge(df_c, pa, by = 'subject_id')
df_m <- merge(df_m, pa, by = 'subject_id')
df_n <- merge(df_n, pa, by = 'subject_id')
df_l <- merge(df_l, pa, by = 'subject_id')


```

First, check association between pa_ee_total and GIMs (covar: age, sex, BMI)

```{r corr-pa-total}

# pa_ee_total is mean total energy expenditure (kcal) per day that waist actical was worn
# Look at the distribution for pa_ee_total

dist_pa_ee_c <- plotDensityHistogram(df_c$pa_ee_total)
dist_pa_ee_l <- plotDensityHistogram(df_l$pa_ee_total) # stool markers (df_c, df_m, df_n) are similar but added df_l because it's the largest distribution 
# Both nearly normal with slight right skew

# Calprotectin

mod_c_pa <- lm(clpn_ln ~ pa_ee_total + age + sex + bmi_final, data = df_c)
summary(mod_c_pa) # slope is positive: 0.0005076
# sig association: p = 0.0380 * 
# age also significant: slope = -0.0076752, p = 0.0346 *

# make the regression plot 

ggplotRegression_c_pa <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = "Mixed Model: Total PA Expenditure vs. Calprotectin (n = 295)",
         subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     ", Intercept = ",signif(fit$coef[[1]],5 ),
                     ", Slope = ",signif(fit$coef[[2]], 5),
                     ", P = ",signif(summary(fit)$coef[2,4], 5)),
       x = "Total PA Expenditure (kcal)",
       y = "Transformed Calprotectin (ln(x + 1))") +
  theme_bw()
}

ggplotRegression_c_pa(mod_c_pa)

c_pa_mx <- ggplotRegression_c_pa(mod_c_pa) 

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/c_pa_mx.jpeg', c_pa_mx, device = 'jpeg', dpi = 300, width = 7, height = 5)



# Myeloperoxidase

mod_m_pa <- lm(mpo_ln ~ pa_ee_total + age + sex + bmi_final, data = df_m)
summary(mod_m_pa) # NS


# make the regression plot 

ggplotRegression_m_pa <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = "Mixed Model: Total PA Expenditure vs. Myeloperoxidase (n = 295)",
         subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     ", Intercept = ",signif(fit$coef[[1]],5 ),
                     ", Slope = ",signif(fit$coef[[2]], 5),
                     ", P = ",signif(summary(fit)$coef[2,4], 5)),
       x = "Total PA Expenditure (kcal)",
       y = "Transformed Myeloperoxidase (ln(x + 3))") +
  theme_bw()
}

ggplotRegression_m_pa(mod_m_pa)

m_pa_mx <- ggplotRegression_m_pa(mod_m_pa) 

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/m_pa_mx.jpeg', m_pa_mx, device = 'jpeg', dpi = 300, width = 7, height = 5)




# Neopterin

mod_n_pa <- lm(nptn_boxcox ~ pa_ee_total + age + sex + bmi_final, data = df_n)
summary(mod_n_pa) # NS
# negative slope is odd

# make the regression plot 

ggplotRegression_n_pa <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = "Mixed Model: Total PA Expenditure vs. Neopterin (n = 289)",
         subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     ", Intercept = ",signif(fit$coef[[1]],5 ),
                     ", Slope = ",signif(fit$coef[[2]], 5),
                     ", P = ",signif(summary(fit)$coef[2,4], 5)),
       x = "Total PA Expenditure (kcal)",
       y = "Transformed Neopterin (Box Cox)") +
  theme_bw()
}

ggplotRegression_n_pa(mod_n_pa)

n_pa_mx <- ggplotRegression_n_pa(mod_n_pa) 

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/n_pa_mx.jpeg', n_pa_mx, device = 'jpeg', dpi = 300, width = 7, height = 5)



# LPS-binding protein (LBP)

mod_l_pa <- lm(lbp_ln ~ pa_ee_total + age + sex + bmi_final, data = df_l)
summary(mod_l_pa) # NS
# negative slope again

# make the regression plot 

ggplotRegression_l_pa <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = "Mixed Model: Total PA Expenditure vs. LBP (n = 348)",
         subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     ", Intercept = ",signif(fit$coef[[1]],5 ),
                     ", Slope = ",signif(fit$coef[[2]], 5),
                     ", P = ",signif(summary(fit)$coef[2,4], 5)),
       x = "Total PA Expenditure (kcal)",
       y = "Transformed LBP (ln(x + 1))") +
  theme_bw()
}

ggplotRegression_l_pa(mod_l_pa)

l_pa_mx <- ggplotRegression_l_pa(mod_l_pa) 

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/l_pa_mx.jpeg', l_pa_mx, device = 'jpeg', dpi = 300, width = 7, height = 5)

# Save in a grid format 

pa_mx <- ggarrange(c_pa_mx, m_pa_mx, n_pa_mx, l_pa_mx,
                      labels = c("A", "B", "C", "D"),
                      nrow = 2, ncol = 2)

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/grid_pa_mx.jpeg', plot = pa_mx, device = 'jpeg', dpi = 300, width = 15, height = 10)

```

Next, let's look at the association between moderate/vigorous PA only (anti-inflammatory compared to light PA) with GIM

```{r add-pa-mod-vig}

# Create a variable (pa_ee_mod_vig) that sums energy expenditure (kcal) from moderate (pa_ee_moderate) and vigorous (pa_ee_vigorous) activity 

df_c$pa_ee_mod_vig <- df_c$pa_ee_moderate + df_c$pa_ee_vigorous
df_c["pa_ee_mod_vig"][is.na(df_c["pa_ee_mod_vig"])] <- 0 # replace NA with 0

df_m$pa_ee_mod_vig <- df_m$pa_ee_moderate + df_m$pa_ee_vigorous
df_m["pa_ee_mod_vig"][is.na(df_m["pa_ee_mod_vig"])] <- 0 # replace NA with 0

df_n$pa_ee_mod_vig <- df_n$pa_ee_moderate + df_n$pa_ee_vigorous
df_n["pa_ee_mod_vig"][is.na(df_n["pa_ee_mod_vig"])] <- 0 # replace NA with 0

df_l$pa_ee_mod_vig <- df_l$pa_ee_moderate + df_l$pa_ee_vigorous
df_l["pa_ee_mod_vig"][is.na(df_l["pa_ee_mod_vig"])] <- 0 # replace NA with 0

```

```{r corr-pa-mod-vig}

# pa_ee_mod_vig is mean total energy expenditure (kcal) from moderate and vigorous activity (excludes light activity) per day that waist actical was worn 

# Look at the distribution for pa_ee_mod_vig
dist_pa_ee_mv_c <- plotDensityHistogram(df_c$pa_ee_mod_vig) # very right skewed, most don't get much mod/vig activity
dist_pa_ee_mv_l <- plotDensityHistogram(df_l$pa_ee_mod_vig) # same shape as _c^

# Calprotectin

mod_c_pa_mv <- lm(clpn_ln ~ pa_ee_mod_vig + age + sex + bmi_final, data = df_c)
summary(mod_c_pa_mv) # NS

# make the regression plot 

ggplotRegression_c_pa_mv <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = "Mixed Model: Moderate & Vigorous PA Expenditure vs. Calprotectin (n = 295)",
         subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     ", Intercept = ",signif(fit$coef[[1]],5 ),
                     ", Slope = ",signif(fit$coef[[2]], 5),
                     ", P = ",signif(summary(fit)$coef[2,4], 5)),
       x = "Moderate & Vigorous PA Expenditure (kcal)",
       y = "Transformed Calprotectin (ln(x + 1))") +
  theme_bw()
}

ggplotRegression_c_pa_mv(mod_c_pa_mv)

c_pa_mv_mx <- ggplotRegression_c_pa_mv(mod_c_pa_mv) 

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/c_pa_mv_mx.jpeg', c_pa_mv_mx, device = 'jpeg', dpi = 300, width = 7, height = 5)



# Myeloperoxidase

mod_m_pa_mv <- lm(mpo_ln ~ pa_ee_mod_vig + age + sex + bmi_final, data = df_m)
summary(mod_m_pa_mv) # NS


# make the regression plot 

ggplotRegression_m_pa_mv <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = "Mixed Model: Moderate & Vigorous PA Expenditure vs. Myeloperoxidase (n = 295)",
         subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     ", Intercept = ",signif(fit$coef[[1]],5 ),
                     ", Slope = ",signif(fit$coef[[2]], 5),
                     ", P = ",signif(summary(fit)$coef[2,4], 5)),
       x = "Moderate & Vigorous PA Expenditure (kcal)",
       y = "Transformed Myeloperoxidase (ln(x + 3))") +
  theme_bw()
}

ggplotRegression_m_pa_mv(mod_m_pa_mv)

m_pa_mv_mx <- ggplotRegression_m_pa_mv(mod_m_pa_mv) 

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/m_pa_mv_mx.jpeg', m_pa_mv_mx, device = 'jpeg', dpi = 300, width = 7, height = 5)




# Neopterin

mod_n_pa_mv <- lm(nptn_boxcox ~ pa_ee_mod_vig + age + sex + bmi_final, data = df_n)
summary(mod_n_pa_mv) # NS


# make the regression plot 

ggplotRegression_n_pa_mv <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = "Mixed Model: Moderate & Vigorous PA Expenditure vs. Neopterin (n = 289)",
         subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     ", Intercept = ",signif(fit$coef[[1]],5 ),
                     ", Slope = ",signif(fit$coef[[2]], 5),
                     ", P = ",signif(summary(fit)$coef[2,4], 5)),
       x = "Moderate & Vigorous PA Expenditure (kcal)",
       y = "Transformed Neopterin (Box Cox)") +
  theme_bw()
}

ggplotRegression_n_pa_mv(mod_n_pa_mv)

n_pa_mv_mx <- ggplotRegression_n_pa_mv(mod_n_pa_mv) 

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/n_pa_mv_mx.jpeg', n_pa_mv_mx, device = 'jpeg', dpi = 300, width = 7, height = 5)



# LPS-binding protein (LBP)

mod_l_pa_mv <- lm(lbp_ln ~ pa_ee_mod_vig + age + sex + bmi_final, data = df_l)
summary(mod_l_pa_mv) # NS


# make the regression plot 

ggplotRegression_l_pa_mv <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue") +
  labs(title = "Mixed Model: Moderate & Vigorous PA Expenditure vs. LBP (n = 348)",
         subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     ", Intercept = ",signif(fit$coef[[1]],5 ),
                     ", Slope = ",signif(fit$coef[[2]], 5),
                     ", P = ",signif(summary(fit)$coef[2,4], 5)),
       x = "Moderate & Vigorous PA Expenditure (kcal)",
       y = "Transformed LBP (ln(x + 1))") +
  theme_bw()
}

ggplotRegression_l_pa_mv(mod_l_pa_mv)

l_pa_mv_mx <- ggplotRegression_l_pa_mv(mod_l_pa_mv) 

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/l_pa_mv_mx.jpeg', l_pa_mv_mx, device = 'jpeg', dpi = 300, width = 7, height = 5)

# Save in a grid format 

pa_mx <- ggarrange(c_pa_mv_mx, m_pa_mv_mx, n_pa_mv_mx, l_pa_mv_mx,
                      labels = c("A", "B", "C", "D"),
                      nrow = 2, ncol = 2)

ggsave(filename = '/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/plots/grid_pa_mv_mx.jpeg', plot = pa_mx, device = 'jpeg', dpi = 300, width = 15, height = 10)


```

**Summary of findings**: 

- pa_ee_total normally distributed for the population, mean 591 kcal, min 158, max 1701 (df_l)
- pa_ee_mod_vig right skewed: mean 191 kcal, min 0, max 960 kcal (df_l)
- pa_ee_total significant positive corr. with calprotectin 
- no associations between pa_ee_mod_vig and any GI inflammation markers 