---
title: "table1_dietguthealth"
author: "Yasmine"
date: "2024-01-12"
output: html_document
---

Running statistics to see associations between vegetable, fruit, saturated fat intake and gut health (Table 2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

```{r import}

df_c <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_c.csv", header = T)

df_m <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_m.csv", header = T)

df_n <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_n.csv", header = T)

df_l <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_l.csv", header = T)

# unfortunately, need to correct structure of some variables each time -->

df_c$subject_id <- as.factor(as.character(df_c$subject_id))
df_c$sex <- df_c$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_c$age_cat <- df_c$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_c$bmi_cat <- df_c$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_m$subject_id <- as.factor(as.character(df_m$subject_id))
df_m$sex <- df_m$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_m$age_cat <- df_m$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_m$bmi_cat <- df_m$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_n$subject_id <- as.factor(as.character(df_n$subject_id))
df_n$sex <- df_n$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_n$age_cat <- df_n$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_n$bmi_cat <- df_n$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_l$subject_id <- as.factor(as.character(df_l$subject_id))
df_l$sex <- df_l$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_l$age_cat <- df_l$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_l$bmi_cat <- df_l$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

```

Models used for analysis: regression of transformed GI inflammation marker ~ dietary component
Covariates: age, sex, BMI

First, vegetables excluding legumes (ASA24) called v_total.x (merged first to df)

```{r asa24-veg-excl-leg}


# Create v_total.x_pk (pk = per 1000 kcal)

df_c$v_total.x_pk <- df_c$v_total.x/(df_c$kcal/1000)
df_m$v_total.x_pk <- df_m$v_total.x/(df_m$kcal/1000)
df_n$v_total.x_pk <- df_n$v_total.x/(df_n$kcal/1000)
df_l$v_total.x_pk <- df_l$v_total.x/(df_l$kcal/1000)


# Calprotectin

mod_v_total.x_pk_c <- lm(clpn_ln ~ v_total.x_pk + age + sex + bmi_final, data = df_c)
summary(mod_v_total.x_pk_c) # NS

# Myeloperoxidase

mod_v_total.x_pk_m <- lm(mpo_ln ~ v_total.x_pk + age + sex + bmi_final, data = df_m)
summary(mod_v_total.x_pk_m) # NS

# Neopterin

mod_v_total.x_pk_n <- lm(nptn_boxcox ~ v_total.x_pk + age + sex + bmi_final, data = df_n)
summary(mod_v_total.x_pk_n) # p = 0.00252 **

# LBP

mod_v_total.x_pk_l <- lm(lbp_ln ~ v_total.x_pk + age + sex + bmi_final, data = df_l)
summary(mod_v_total.x_pk_l) # NS


```
Next, vegetables excluding legumes (FFQ) called v_total.y (merged second to df)

```{r ffq-veg-excl-leg}

# Create v_total.y_pk (pk = per 1000 kcal)

df_c$v_total.y_pk <- df_c$v_total.y/(df_c$kcal/1000)
df_m$v_total.y_pk <- df_m$v_total.y/(df_m$kcal/1000)
df_n$v_total.y_pk <- df_n$v_total.y/(df_n$kcal/1000)
df_l$v_total.y_pk <- df_l$v_total.y/(df_l$kcal/1000)

# Calprotectin

mod_v_total.y_pk_c <- lm(clpn_ln ~ v_total.y_pk + age + sex + bmi_final, data = df_c)
summary(mod_v_total.y_pk_c) # NS

# Myeloperoxidase

mod_v_total.y_pk_m <- lm(mpo_ln ~ v_total.y_pk + age + sex + bmi_final, data = df_m)
summary(mod_v_total.y_pk_m) # NS

# Neopterin

mod_v_total.y_pk_n <- lm(nptn_boxcox ~ v_total.y_pk + age + sex + bmi_final, data = df_n)
summary(mod_v_total.y_pk_n) # p = 0.00252 **

# LBP

mod_v_total.y_pk_l <- lm(lbp_ln ~ v_total.y_pk + age + sex + bmi_final, data = df_l)
summary(mod_v_total.y_pk_l) # NS

```

Now, legumes (ASA24) called v_legumes.x (merged first to df)

```{r asa24-legumes}

# Create v_legumes.x_pk (pk = per 1000 kcal)

df_c$v_legumes.x_pk <- df_c$v_legumes.x/(df_c$kcal/1000)
df_m$v_legumes.x_pk <- df_m$v_legumes.x/(df_m$kcal/1000)
df_n$v_legumes.x_pk <- df_n$v_legumes.x/(df_n$kcal/1000)
df_l$v_legumes.x_pk <- df_l$v_legumes.x/(df_l$kcal/1000)

# Calprotectin

mod_v_legumes.x_pk_c <- lm(clpn_ln ~ v_legumes.x_pk + age + sex + bmi_final, data = df_c)
summary(mod_v_legumes.x_pk_c) # NS

# Myeloperoxidase

mod_v_legumes.x_pk_m <- lm(mpo_ln ~ v_legumes.x_pk + age + sex + bmi_final, data = df_m)
summary(mod_v_legumes.x_pk_m) # NS

# Neopterin

mod_v_legumes.x_pk_n <- lm(nptn_boxcox ~ v_legumes.x_pk + age + sex + bmi_final, data = df_n)
summary(mod_v_legumes.x_pk_n) # 

# LBP

mod_v_legumes.x_pk_l <- lm(lbp_ln ~ v_legumes.x_pk + age + sex + bmi_final, data = df_l)
summary(mod_v_legumes.x_pk_l) # NS


```
Next, legumes (FFQ) called v_legumes.y (merged second to df)

```{r ffq-legumes}


# Create v_legumes.y_pk (pk = per 1000 kcal)

df_c$v_legumes.y_pk <- df_c$v_legumes.y/(df_c$kcal/1000)
df_m$v_legumes.y_pk <- df_m$v_legumes.y/(df_m$kcal/1000)
df_n$v_legumes.y_pk <- df_n$v_legumes.y/(df_n$kcal/1000)
df_l$v_legumes.y_pk <- df_l$v_legumes.y/(df_l$kcal/1000)

# Calprotectin

mod_v_legumes.y_pk_c <- lm(clpn_ln ~ v_legumes.y_pk + age + sex + bmi_final, data = df_c)
summary(mod_v_legumes.y_pk_c) # NS

# Myeloperoxidase

mod_v_legumes.y_pk_m <- lm(mpo_ln ~ v_legumes.y_pk + age + sex + bmi_final, data = df_m)
summary(mod_v_legumes.y_pk_m) # NS

# Neopterin

mod_v_legumes.y_pk_n <- lm(nptn_boxcox ~ v_legumes.y_pk + age + sex + bmi_final, data = df_n)
summary(mod_v_legumes.y_pk_n) # p = 0.00252 **

# LBP

mod_v_legumes.y_pk_l <- lm(lbp_ln ~ v_legumes.y_pk + age + sex + bmi_final, data = df_l)
summary(mod_v_legumes.y_pk_l) # NS

```
Up next, true total vegetables (v_total + v_legumes) = v_t_total.x 

```{r asa24-veg-t_tot}

# Create v_t_total.x_pk (pk = per 1000 kcal)

df_c$v_t_total.x_pk <- df_c$v_t_total.x/(df_c$kcal/1000)
df_m$v_t_total.x_pk <- df_m$v_t_total.x/(df_m$kcal/1000)
df_n$v_t_total.x_pk <- df_n$v_t_total.x/(df_n$kcal/1000)
df_l$v_t_total.x_pk <- df_l$v_t_total.x/(df_l$kcal/1000)


# Calprotectin

mod_v_t_total.x_pk_c <- lm(clpn_ln ~ v_t_total.x_pk + age + sex + bmi_final, data = df_c)
summary(mod_v_t_total.x_pk_c) # NS

# Myeloperoxidase

mod_v_t_total.x_pk_m <- lm(mpo_ln ~ v_t_total.x_pk + age + sex + bmi_final, data = df_m)
summary(mod_v_t_total.x_pk_m) # NS

# Neopterin

mod_v_t_total.x_pk_n <- lm(nptn_boxcox ~ v_t_total.x_pk + age + sex + bmi_final, data = df_n)
summary(mod_v_t_total.x_pk_n) # p = 0.01919 * 

# LBP

mod_v_t_total.x_pk_l <- lm(lbp_ln ~ v_t_total.x_pk + age + sex + bmi_final, data = df_l)
summary(mod_v_t_total.x_pk_l) # NS


```
Next, true total vegetables (FFQ) called v_t_total.y (merged second to df)

```{r ffq-veg-t-tot}

# Create v_t_total.y_pk (pk = per 1000 kcal)

df_c$v_t_total.y_pk <- df_c$v_t_total.y/(df_c$kcal/1000)
df_m$v_t_total.y_pk <- df_m$v_t_total.y/(df_m$kcal/1000)
df_n$v_t_total.y_pk <- df_n$v_t_total.y/(df_n$kcal/1000)
df_l$v_t_total.y_pk <- df_l$v_t_total.y/(df_l$kcal/1000)


# Calprotectin

mod_v_t_total.y_pk_c <- lm(clpn_ln ~ v_t_total.y_pk + age + sex + bmi_final, data = df_c)
summary(mod_v_t_total.y_pk_c) # NS

# Myeloperoxidase

mod_v_t_total.y_pk_m <- lm(mpo_ln ~ v_t_total.y_pk + age + sex + bmi_final, data = df_m)
summary(mod_v_t_total.y_pk_m) # NS

# Neopterin

mod_v_t_total.y_pk_n <- lm(nptn_boxcox ~ v_t_total.y_pk + age + sex + bmi_final, data = df_n)
summary(mod_v_t_total.y_pk_n) # p = 0.00252 **

# LBP

mod_v_t_total.y_pk_l <- lm(lbp_ln ~ v_t_total.y_pk + age + sex + bmi_final, data = df_l)
summary(mod_v_t_total.y_pk_l) # NS

```
Let's look at total fruit (f_total)

```{r asa24-f-tot}

# Create f_total.x_pk (pk = per 1000 kcal)

df_c$f_total.x_pk <- df_c$f_total.x/(df_c$kcal/1000)
df_m$f_total.x_pk <- df_m$f_total.x/(df_m$kcal/1000)
df_n$f_total.x_pk <- df_n$f_total.x/(df_n$kcal/1000)
df_l$f_total.x_pk <- df_l$f_total.x/(df_l$kcal/1000)


# Calprotectin

mod_f_total.x_pk_c <- lm(clpn_ln ~ f_total.x_pk + age + sex + bmi_final, data = df_c)
summary(mod_f_total.x_pk_c) # NS

# Myeloperoxidase

mod_f_total.x_pk_m <- lm(mpo_ln ~ f_total.x_pk + age + sex + bmi_final, data = df_m)
summary(mod_f_total.x_pk_m) # NS

# Neopterin

mod_f_total.x_pk_n <- lm(nptn_boxcox ~ f_total.x_pk + age + sex + bmi_final, data = df_n)
summary(mod_f_total.x_pk_n) # p =  0.02475 * 

# LBP

mod_f_total.x_pk_l <- lm(lbp_ln ~ f_total.x_pk + age + sex + bmi_final, data = df_l)
summary(mod_f_total.x_pk_l) # NS


```
Next, total fruit (FFQ) called f_total.y (merged second to df)

```{r ffq-f-tot}

# Create f_total.y_pk (pk = per 1000 kcal)

df_c$f_total.y_pk <- df_c$f_total.y/(df_c$kcal/1000)
df_m$f_total.y_pk <- df_m$f_total.y/(df_m$kcal/1000)
df_n$f_total.y_pk <- df_n$f_total.y/(df_n$kcal/1000)
df_l$f_total.y_pk <- df_l$f_total.y/(df_l$kcal/1000)


# Calprotectin

mod_f_total.y_pk_c <- lm(clpn_ln ~ f_total.y_pk + age + sex + bmi_final, data = df_c)
summary(mod_f_total.y_pk_c) # NS

# Myeloperoxidase

mod_f_total.y_pk_m <- lm(mpo_ln ~ f_total.y_pk + age + sex + bmi_final, data = df_m)
summary(mod_f_total.y_pk_m) # NS

# Neopterin

mod_f_total.y_pk_n <- lm(nptn_boxcox ~ f_total.y_pk + age + sex + bmi_final, data = df_n)
summary(mod_f_total.y_pk_n) # p = 0.00252 **

# LBP

mod_f_total.y_pk_l <- lm(lbp_ln ~ f_total.y_pk + age + sex + bmi_final, data = df_l)
summary(mod_f_total.y_pk_l) # NS

```
Finally, looking at saturated fat (sfat)

```{r asa24-sfat}

# Create sfat_pk (pk = per 1000 kcal)

df_c$sfat_pk <- df_c$sfat/(df_c$kcal/1000)
df_m$sfat_pk <- df_m$sfat/(df_m$kcal/1000)
df_n$sfat_pk <- df_n$sfat/(df_n$kcal/1000)
df_l$sfat_pk <- df_l$sfat/(df_l$kcal/1000)

# Calprotectin

mod_sfat_pk_c <- lm(clpn_ln ~ sfat_pk + age + sex + bmi_final, data = df_c)
summary(mod_sfat_pk_c) # NS

# Myeloperoxidase

mod_sfat_pk_m <- lm(mpo_ln ~ sfat_pk + age + sex + bmi_final, data = df_m)
summary(mod_sfat_pk_m) # NS

# Neopterin

mod_sfat_pk_n <- lm(nptn_boxcox ~ sfat_pk + age + sex + bmi_final, data = df_n)
summary(mod_sfat_pk_n) # p =  0.02475 * 

# LBP

mod_sfat_pk_l <- lm(lbp_ln ~ sfat_pk + age + sex + bmi_final, data = df_l)
summary(mod_sfat_pk_l) # NS


```
Last, saturated fat (FFQ) called dt_sfat

```{r ffq-sfat}

# Create dt_sfat_pk (pk = per 1000 kcal)

df_c$dt_sfat_pk <- df_c$dt_sfat/(df_c$kcal/1000)
df_m$dt_sfat_pk <- df_m$dt_sfat/(df_m$kcal/1000)
df_n$dt_sfat_pk <- df_n$dt_sfat/(df_n$kcal/1000)
df_l$dt_sfat_pk <- df_l$dt_sfat/(df_l$kcal/1000)

# Calprotectin

mod_dt_sfat_pk_c <- lm(clpn_ln ~ dt_sfat_pk + age + sex + bmi_final, data = df_c)
summary(mod_dt_sfat_pk_c) # NS

# Myeloperoxidase

mod_dt_sfat_pk_m <- lm(mpo_ln ~ dt_sfat_pk + age + sex + bmi_final, data = df_m)
summary(mod_dt_sfat_pk_m) # NS

# Neopterin

mod_dt_sfat_pk_n <- lm(nptn_boxcox ~ dt_sfat_pk + age + sex + bmi_final, data = df_n)
summary(mod_dt_sfat_pk_n) # 

# LBP

mod_dt_sfat_pk_l <- lm(lbp_ln ~ dt_sfat_pk + age + sex + bmi_final, data = df_l)
summary(mod_dt_sfat_pk_l) # NS

```

```

**Summary of Findings**: 

- Recent intake of vegetables excluding legumes, total vegetables, and total fruit negatively correlated with neopterin levels 