---
title: "table1_dietguthealth"
author: "Yasmine"
date: "2024-01-12"
output: html_document
---

Running statistics to see associations between fiber and subclinical calprotectin and MPO (Table 1) per 1000 kcal

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

```{r import}

df_c <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_c.csv", header = T)

df_m <- read.csv("/Users/yasmine.bouzid/Documents/GitHub/diet_guthealth/data/df_m.csv", header = T)

# For both, we're excluding those with calprotectin and mpo levels above clinical threshold:


# Label based off the clinical designation levels

df_c$clin_level_clpn[df_c$Stool.calprotectin < 50] <- "negative"

df_c$clin_level_clpn[df_c$Stool.calprotectin >= 50 & df_c$Stool.calprotectin <= 100] <- "bord_positive"

df_c$clin_level_clpn[df_c$Stool.calprotectin > 100] <- "positive"

# Convert variable type to factor, and order them 
df_c$clin_level_clpn <- factor(df_c$clin_level_clpn, ordered = T,
                                  c("negative", "bord_positive", "positive"))

table(df_c$clin_level_clpn)

df_c_s <- df_c %>% filter(clin_level_clpn != "positive") # n = 253, check. 

df_c <- df_c_s
rm(df_c_s)


# Label based off the clinical designation levels

df_m$clin_level_mpo[df_m$fecal_mpo < 2000] <- "negative"

df_m$clin_level_mpo[df_m$fecal_mpo >= 2000] <- "positive"

# Convert variable type to factor, and order them 
df_m$clin_level_mpo <- factor(df_m$clin_level_mpo, ordered = T,
                                  c("negative", "positive"))

table(df_m$clin_level_mpo)

df_m_s <- df_m %>% filter(clin_level_mpo != "positive") # n = 286, check.

df_m <- df_m_s
rm(df_m_s)


```


```{r correct-class}

# unfortunately, need to correct structure of some variables each time -->

df_c$subject_id <- as.factor(as.character(df_c$subject_id))
df_c$sex <- df_c$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_c$age_cat <- df_c$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_c$bmi_cat <- df_c$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

df_m$subject_id <- as.factor(as.character(df_m$subject_id))
df_m$sex <- df_m$sex %>% recode_factor('1' = 'Male', '2' = "Female")
df_m$age_cat <- df_m$age_cat %>% recode_factor('1' = '18-33.99 years', '2' = "34-49.99 years", '3' = "50-65 years")
df_m$bmi_cat <- df_m$bmi_cat %>% recode_factor('1' = 'Normal', '2' = "Overweight", '3' = "Obese")

```





Models used for analysis: regression of transformed GI inflammation marker ~ dietary component
Covariates: age, sex, BMI

```{r asa24-total-fiber}

# Create fibe_pk (pk = per 1000 kcal)

df_c$fibe_pk <- df_c$fibe/(df_c$kcal/1000)
df_m$fibe_pk <- df_m$fibe/(df_m$kcal/1000)

# Calprotectin

mod_fibe_pk_c <- lm(clpn_ln ~ fibe_pk + age + sex + bmi_final, data = df_c)
summary(mod_fibe_pk_c) 

# Myeloperoxidase

mod_fibe_pk_m <- lm(mpo_ln ~ fibe_pk + age + sex + bmi_final, data = df_m)
summary(mod_fibe_pk_m) 


```
Next, we'll look at total fiber from FFQ (dt_fibe)

```{r ffq-total-fiber}

# Create dt_fibe_pk (pk = per 1000 kcal)

df_c$dt_fibe_pk <- df_c$dt_fibe/(df_c$kcal/1000)
df_m$dt_fibe_pk <- df_m$dt_fibe/(df_m$kcal/1000)

# Calprotectin

mod_dt_fibe_pk_c <- lm(clpn_ln ~ dt_fibe_pk + age + sex + bmi_final, data = df_c)
summary(mod_dt_fibe_pk_c) # P = 0.0285 *  

# Myeloperoxidase

mod_dt_fibe_pk_m <- lm(mpo_ln ~ dt_fibe_pk + age + sex + bmi_final, data = df_m)
summary(mod_dt_fibe_pk_m) # NS


```

Last, we'll look at soluble fiber from the FFQ (dt_fiber_sol)

```{r ffq-sol-fiber}

# Create dt_fiber_sol_pk (pk = per 1000 kcal)

df_c$dt_fiber_sol_pk <- df_c$dt_fiber_sol/(df_c$kcal/1000)
df_m$dt_fiber_sol_pk <- df_m$dt_fiber_sol/(df_m$kcal/1000)

# Calprotectin

mod_dt_fiber_sol_pk_c <- lm(clpn_ln ~ dt_fiber_sol_pk + age + sex + bmi_final, data = df_c)
summary(mod_dt_fiber_sol_pk_c) # NS

# Myeloperoxidase

mod_dt_fiber_sol_pk_m <- lm(mpo_ln ~ dt_fiber_sol_pk + age + sex + bmi_final, data = df_m)
summary(mod_dt_fiber_sol_pk_m) # NS


```

**Summary of Findings**: 

- Habitual total fiber intake negatively correlated with subclinical calprotectin